---
title: "Apportionment simulations"
author: "Kari Fenske, Curry Cunningham, Dana Hanselman"
date: "August 16, 2019"
output: html_document
---


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#install the libraries
require(dplyr)
require(reshape2)
require(tidyverse)
require(ggplot2)
#for extract-pars.R
#require(readxl)
require(xlsx)
# for sample-age-comps.R
require(gtools) 
#for .dat file building
require(PBSmodelling)
require(R2admb)

###THESE ITEMS NEED MANUAL INPUT!!!###
n.apports <- 10 #number of apportionment options that have been run and are to be compiled here
apports <- c(1,2,3,4,5,6,7,8,9,10) #the folder numbers of the apportionment options to be compiled
#"Equal-1","Fixed-2","Equilibrium-3","NPFMC-4","Exp_survey_wt-5","Exp_fishery_wt-6","Non-Exp_NPFMC-7",
#"Partial_fixed-8","Age_based-9","RandEff-10"","All_to_one-11","TermYear_biom-12","Penalized-13","Non-exp_Agebased-14"
apport.names <- list("Equal","Fixed","Equilibrium","NPFMC","Exp_survey_wt","Exp_fishery_wt","Non-Exp_NPFMC",
"Partial_fixed","Age_based","All_to_one")#,"Equilibrium","NPFMC") #list the names of the apportionment options being compared


#read in the apportionment R object files for indexing
wd <- ("C:/Repositories/Sablefish_ApportionmentStrategies") #may need to manually enter your wd here
dir.output <- file.path(wd,"output/")
dir.admb <- file.path(wd,"admb")

#setwd(dir.output)
n.year <- readRDS(file.path(wd,"output","n.year.rds"))
years <- readRDS(file.path(wd,"output","years.rds"))
n.sims <- readRDS(file.path(wd,"output","n.sims.rds"))
sims <- readRDS(file.path(wd,"output","sims.rds"))
n.age <- readRDS(file.path(wd,"output","n.age.rds"))
ages <- readRDS(file.path(wd,"output","ages.rds"))
n.sex <- readRDS(file.path(wd,"output","n.sex.rds"))
sexes <- readRDS(file.path(wd,"output","sexes.rds"))
n.area <- readRDS(file.path(wd,"output","n.area.rds"))
areas <- readRDS(file.path(wd,"output","areas.rds"))
age.rec <- readRDS(file.path(wd,"output","age.rec.rds"))
A <- readRDS(file.path(wd,"output","A.rds"))
apport.opt <- readRDS(file.path(wd,"output","apport.opt.rds"))
n.fish <- readRDS(file.path(wd,"output","n.fish.rds"))
fish <- readRDS(file.path(wd,"output","fish.rds"))
n.surv <- readRDS(file.path(wd,"output","n.surv.rds"))
surv.name <- readRDS(file.path(wd,"output","surv.name.rds"))

#plotting things
forproj.styr <- 44 #define the year (y) where EM is first run 
OMyears <- c(1977:2018) #conditioning period years for OM
rpn_years <- c(15:43) #LL survey index years
rpw_years <- c(15:42) #fixed gear fishery years
mgmt_rep_years <- c(1960:2018) # these are the years for management EM data (for comparison plotting)

#read in data from single area management model/Kari's single model (it's sort of hybrid data)
mgmt_dat <- readList(file.path(dir.admb,"Single_area/permanant_tem_single2018.dat"))
#read in data from the single area mgmt report file
mgmt_rep <- readList(file.path(dir.admb,"Management/tem_simplified.rep"))
#read in data from the conditioning OM years
OM <- readList(file.path(dir.admb,"Single_area/tem_single2018.dat"))

#read in OM objects as you use them below
#read in EM objects as you use them below

```

# General information
Sablefish apportionment of ABC to management regions has been fixed at 2013 ratios since 2014. An objective of the analyses presented here is to examine pros and cons of a suite of sablefish ABC apportionment methods. This document presents the apportionment simulation work to date. This is a work in progress and we are seeking feedback on OM and EM set up, alternative performance metrics that would be useful in making apportionment method recommendations, and any other suggestions for simulation and model performance. Bonus points if you can suggest better ways to graph things.

These apportionment simulation analyses contain two primary components, a 6-area operating model (OM) and a 1-area estimation model (EM). The OM is spatial so that potential spatial dynamics in fleet or fish behavior (via catchability, selectivity, and fish movement) may be simulated. The OM simulates data in two periods - a deterministic conditioning period which occurs for years 1977-2018 and is the same across simulations, and a stochastic forward projecting portion which runs for years 2019-`r 1976+n.year-1`. The EM is similar to the EM used for sablefish management, but begins in 1977 instead of 1960, does not include length compositions, and does not fit a trawl survey. After the conditioning period data is generated in the OM, the OM population is sampled, and simulated data from the OM is combined into a single area dataset which is passed to the EM.  

In the forward projecting period, the OM-EM is iterative, looping through years. The order of operations for year y are: OM - Read in previous year's apportioned ABC by area, estimate the F required to catch ABC for each area, apply F and M to OM population, move fish between areas, sample the population for fishery and longline survey abundance indices and for longline survey and fixed gear fishery age compositions, build the .dat file and pass it to the EM in ADMB, run the EM and get an estimate of the next year's ABC (then repeat for the next year).

At present for each apportionment method we examine, we are running `r n.sims` simulations covering years 1977-`r 1976+n.year-1`. For November, we hope to present results for 100 (or more) simulated datasets each looping over years 2019-2049 (30 years for forward looping years).

There are some inherent assumptions in this simulation we would like to present up front:
1) we assume ABC=TAC and 100% of apportioned ABC is caught in each region
2) we do not correct for whale depredation in the ABC or survey index
3) recruitment occurs at age 2 and recruitment is split equally between males and females


# Alternative apportionment scenarios 
In the analyses presented in this document we examine X apportion methods: `r apport.names`.  
The future plan is to examine this suite:  
1 - Equal (Each region receives 1/6 of the ABC)  
2 - Fixed (The apportionment proportions from the 2013 assessment that have been applied as fixed proportions for 2014-2018)   
3 - Equilibrium (Based on the stationary distribution of the movement rates)  
4 - NPFMC (5-yr exponentially weighted moving average of fishery and survey indices; survey weight is 2x fishery weight)  
5 - Exp_survey_wt (Similar to ‘NPFMC’ option but using survey index only)  
6 - Exp_fishery_wt (Similar to ‘NPFMC’ option but using fishery index only)   
7 - Non-Exp_NPFMC (A 5-yr moving average of fishery and survey indices)  
8 - Partial_fixed (BS and AI receive 10%  of the ABC each, WG, CG, WY, and EY are apportioned based on NPFMC method)   
9 - Age_based (Based on the proportions of fish at age of 50% maturity in each area - i.e. areas with greater proportion of fish at age of 50% maturity or greater will be apportioned a greater proportion of ABC)  
10 - (needs to be added) Terminal year of longline survey (no exp wts), proxy for RE model.  Maybe sub this for equilibrium movement?
``` {r, echo=FALSE}
#these are a few apportionment options we've considered and may do, but either are for internal (research) use, aren't a priority, or aren't yet functioning. 
#10 - RandEff (Apportionment to region based on the proportions of biomass estimated by the RE model applied to the longline survey, using 0.05 CV)    
#11 - All_to_one (all apportionment to one area - for research purposes of showing extremes ONLY)  
#12 - TermYear_biom (Based on terminal year survey biomass)  
#13 - Penalized 
#14 - Non-exp_Agebased  
```

# Conditioning period
The OM model set up begins by establishing initial numbers at age for each area in 1976. Initial N for 1976 is input as the 2018 management EM totaly abundance estimated for 1977. This is split into `r n.area` spatial areas using the proportion of abundance by area from the longline survey abundance estimate and split into initial age and sex proportions using proportion by sex and age from 2018 management EM for 1976 numbers at age.  Numbers at age are conveted to biomass at age using the age-weight relationship as described in the sablefish 2018 SAFE report.

OM conditioning period is deterministic and the same for all simulation iterations. Movement is specified between `r n.area` and the OM is set up to accomodate age-based movement at present movement is un-varying over ages. The OM conditioning period is deterministic, with recruitment from the EM read in, catch, read in and the F which generated that catch estimated and applied to the population. 

Movement rates used in the OM are externally generated and based on the Hanselman et al. 2015 "Move it or lose it..." paper. The model described in the paper was re-run for `r n.area' and those movement values are input.
```{r, echo=FALSE}
movement_sample <- matrix(data=NA, nrow=n.area,ncol=n.area,dimnames=list(c("From EY","From WY","From CG","From WG","From BS","From AI"),c("To EY","To WY","To CG","To WG","To BS","To AI")))

movement_sample[1,] <- c(0.744367,0.076350,0.149227,0.026232,0.002098,0.001726)
movement_sample[2,] <- c(0.137189,0.191575,0.478710,0.152110,0.021954,0.018462)
movement_sample[3,] <- c(0.108443,0.193604,0.486216,0.162973,0.026394,0.022370)
movement_sample[4,] <- c(0.037484,0.120964,0.320461,0.290733,0.122029,0.108328)
movement_sample[5,] <- c(0.005357,0.031088,0.092290,0.215184,0.627622,0.028460)
movement_sample[6,] <- c(0.002495,0.014990,0.045060,0.113111,0.046442,0.777902)
movement_sample
```
  
See the Appendix for OM equations.  

# Conditioning period validation
The figures and tables below are presented to show the ability of the spatial (with movement) OM conditioning period to match the current EM historical data and to further describe the methods for setting up the OM. In general, the conditioning period matches very well to the EM. However some differences exist due to movement and spatial OM parameters which differ from the management EM. Because this isn't a full MSE, alternative states of nature and alternative realities are not the focus. 

Abundance in numbers (millions of fish, summed over areas) generated from the OM matches the Management model EM.
```{r, echo=FALSE}
dir.temp <- paste0(dir.output,"/Apport.Option_",1) #OM for conditioning period is the same across apportionment options, so pulling conditioning period output from folder 1
N <- readRDS(file.path(dir.temp,"N.rds")) #OM conditioning period N 
B <- readRDS(file.path(dir.temp,"B.rds")) 
ssb <- readRDS(file.path(dir.temp,"ssb.rds"))
temp.catchnumbiom <- readRDS(file.path(dir.temp,"temp.catchnumbiom.rds"))
C.b <- readRDS(file.path(dir.temp,"C.b.rds"))
harvest.b <- readRDS(file.path(dir.temp,"harvest.b.rds"))
cond.rec <- readRDS(file.path(dir.temp,"cond.Recruitment.rds"))
recruits.area <- readRDS(file.path(dir.temp,"recruits.area.rds"))
va <- readRDS(file.path(dir.temp,"va.rds"))
va_surv <- readRDS(file.path(dir.temp,"va_surv.rds"))
```

``` {r, echo=FALSE}
#management N, by year (summed across ages, sexes) for 1960-2018 is:
mgmtN <- c(227.569424,225.133078,214.348775,202.312974,234.635297,228.675439,
           223.024404,233.474335,226.238316,210.835746,197.26262,183.638328,
           162.608407,138.835745,123.192946,118.53197,108.080044,93.414099,
           83.964022,159.561953,166.4444777,157.2009806,180.5002287,183.5209969,
           205.8722927,184.1836206,181.7986075,174.4102706,150.830862,129.4616883,
           113.7451334,121.4683674,104.0961752,111.7859086,99.2854377,89.0569055,
           82.837678,87.163893,77.2629253,96.086156,100.0417057,96.0967192,
           124.029391,113.7307861,110.2138382,99.9701582,95.7489713,89.5424044,
           84.6328623,80.147036,86.599641,79.9309313,77.8211843,67.23660782,
           64.6833935,68.2611382,208.8285434,225.8192621,212.4078021)

#OM conditioning period N, by year (summed across ages, sexes, areas)
simN_sum <- apply(N[,,,,1],2,sum) # sim 1 results only (sims are all identical for conditioning period)
par(mfrow=c(1,1))
plot(mgmtN~mgmt_rep_years, ylim=c(0,250),typ="l",lwd=3,col="red",xlab="Year",ylab="Abundance (millions)")
lines(simN_sum[2:43]~OMyears,typ="l",lwd=3,col="black") 
legend("bottomleft", legend=c("Management EM", "Conditioning OM"),lty=c(1,1),lwd=c(3,3), col=c("red","black"))

```


Biomass (kt, summed over areas) generated from the OM matches the Management model EM.
```{r, echo=FALSE}
#mgmt_rep$Tot_biom #management EM total biomass
simB_sum <- apply(B[,,,,1],2,sum) #OM conditioning period B, by year (summed across ages, sexes, areas), from sim 1
par(mfrow=c(1,1))
plot(mgmt_rep$Tot_biom~mgmt_rep_years, ylim=c(0,650),typ="l",lwd=3,col="red",xlab="Year",ylab="Biomass (kt)")
lines(simB_sum[2:43]~OMyears,typ="l",lwd=3,col="black") 
legend("bottomleft", legend=c("Management EM", "Conditioning OM"),lty=c(1,1),lwd=c(3,3), col=c("red","black"))

```


Spawning biomass (kt, summed over areas) generated from the OM matches the Management model EM.
```{r, echo=FALSE}
#mgmt_rep$SpBiom #mgmt EM ssb
simssb_sum <- apply(ssb[,,,,1],3,sum) #OM conditioning period ssb, summed across ages, sexes, areas
par(mfrow=c(1,1))
plot(mgmt_rep$SpBiom~mgmt_rep_years, ylim=c(0,300),typ="l",lwd=3,col="red",xlab="Year",ylab="Spawning Biomass (kt)")
lines(simssb_sum[2:43]~OMyears,typ="l",lwd=3,col="black") 
legend("bottomleft", legend=c("Management EM", "Conditioning OM"),lty=c(1,1),lwd=c(3,3), col=c("red","black"))

```


OM catch was designed to match EM observed catch. An F-solving function in the OM takes input catch and estimates the fishing mortality required to take that catch, using OM numbers at age and selectivity. The estimated F rate is then used in the generating abundance for the OM.
```{r, echo=FALSE}
mgmtCatch <- mgmt_rep$Obs_CatchTrawl + mgmt_rep$Obs_CatchFixedGear #total catch from mgmt EM
read.in.catch <- apply(temp.catchnumbiom,1,sum)
sim_C.b_sum <- apply(C.b[,,,,1],2,sum)
sim_harv_sum <- apply(harvest.b[,,,,,1],2,sum)

#Option 1
#plot(mgmtCatch~mgmt_rep_years, ylim=c(0,55),typ="l",lwd=3,col="red",xlab="Year",ylab="Catch (kt)")
#lines(read.in.catch[2:(length(OMyears)+1)]~OMyears,typ="l",lwd=3,col="black")
#lines(sim_C.b_sum[2:(length(OMyears)+1)]~OMyears,typ="l",lwd=3,col="blue")
#lines(sim_harv_sum[2:(length(OMyears)+1)]~OMyears,typ="l",lwd=3,col="blue")
#legend("topright", legend=c("Management EM Catch", "Conditioning OM Input Catch", "Conditioning OM Est. Catch", "Conditioning OM Est. Harvest"),lty=c(1,1,1,1),lwd=c(3,3,3,3), col=c("red","black","blue","blue"))
#option 2 - without harvest line
plot(mgmtCatch~mgmt_rep_years, ylim=c(0,55),typ="l",lwd=3,col="red",xlab="Year",ylab="Catch (kt)")
lines(read.in.catch[2:(length(OMyears)+1)]~OMyears,typ="l",lwd=3,col="black")
lines(sim_C.b_sum[2:(length(OMyears)+1)]~OMyears,typ="l",lwd=3,col="blue")
legend("topright", legend=c("Management EM Catch", "Conditioning OM Input Catch", "Conditioning OM Est. Catch"),lty=c(1,1,1),lwd=c(3,3,3), col=c("red","black","blue"))

```


OM recruitment was designed to match EM recruitment.  Mgmt EM recruitment values (in numbers) for 1977-2018 are read in and split into `r n.area` areas based on the proportion of age-2 sablefish in the longline survey in each area, and split equally between males and females.  
**N**(h,y,a,m,i) = Rec.Proportion(m) x 0.5 x **Recruitment**(y-1),   
where h, y, a, m, and i are subscripts for sex, year, age, area, and simulation, respectively. This figure shows management EM recruitment and OM recruitment (summed over areas).
``` {r, echo=FALSE}
mgmt_rep_recruitment <- c(122.9618,20.0246,15.45504,16.50548,57.0088,18.69596,18.98642,36.665,21.0668,15.71956,
       17.67248,16.18402,9.21424,7.10778,8.0244,16.61306,9.04232,4.41974,5.29922,86.386,25.42,9.9051,
       42.303,23.9728,43.5138,2.33748,19.53042,18.80922,3.6228,4.52162,6.87626,27.2682,1.286332,23.7892,4.59836,
       5.31364,7.73184,16.88416,2.30162,29.923,16.93562,10.15266,41.4692,6.34222,12.67834,6.05714,10.71728,
       7.79438,8.44232,7.82108,17.84292,5.03906,9.29366,1.01781,7.98824,13.20866,150.343,40.1198,12.4912)
#cond.rec$Recruitment #recruits for the conditioning period - these values are the input to the OM, once split into areas

plot(mgmt_rep_recruitment~mgmt_rep_years,xlim=c(1960,2020),typ="l",lwd=3,col="red",ylim=c(0,175))
lines(cond.rec~OMyears,lty=3,lwd=3,col="black")
legend("topleft", legend=c("Management EM Recruitment", "Conditioning OM Recruitment"),lty=c(1,3),lwd=c(3,3), col=c("red","black"))

```

The values for age-2 proportions by area (for order BS, AI, WG, CG, WY, EY/SEO) from the longline surevey are: 
``` {r, echo=FALSE}
#conditioning recruitment by area
rec.by.area.props <- c(0.14,0.07,0.14,0.43,0.14,0.09)
```
`r rec.by.area.props`.

``` {r, echo=FALSE}
sum_c.rec1 <- apply(N[,2:43,1,,1],c(2,3),sum) #recruitment by year and area
sum_c.rec2 <- apply(N[,2:43,1,,1],2,sum) #recruitment by year
plot(sum_c.rec1[,1]~OMyears,typ="l",lwd=4,col="black",ylim=c(0,55),ylab="Recruitment @ age 2 (millions)", xlab="Year",lty=1)
lines(sum_c.rec1[,2]~OMyears,typ="l",lwd=4,col="green",lty=2)
lines(sum_c.rec1[,3]~OMyears,typ="l",lwd=4,col="red",lty=1)
lines(sum_c.rec1[,4]~OMyears,typ="l",lwd=4,col="grey",lty=1)
lines(sum_c.rec1[,5]~OMyears,typ="l",lwd=4,col="orange",lty=3)
lines(sum_c.rec1[,6]~OMyears,typ="l",lwd=4,col="blue",lty=3)
legend("topleft", legend=c("BS", "AI","WG", "CG","WY", "EY/SEO"),lty=c(1,2,1,1,3,3),lwd=c(3,3,3,3,3,3), col=c("black","green","red","grey","orange","blue"))

```

Recruitment for the forward projecting period is the same across apportionment simulations and does not have a stock recruitment relationship. Recruitment for n.years x n.sims is drawn once (from a multinomial distribution) and used for all apportionment methods. Mean recruitment (average log-recruitment) and recruitment standard deviation for the multinomial is from the management EM. Recruitment in each year is divided into OM spatial areas based on the mean proportion of age-2 (recruitment age) sablefish in each area from the longline survey for 1977-2018.

For comparison, the conditioning period recruitment draws for `r n.sims` compared to management EM estimated recruitment is shown here, though note that these values are NOT USED in the conditioning period. 
```{r, echo=FALSE}
sum.rec.area <- apply(recruits.area,c(1,3),sum)
#mgmt_rep_recruitment
plot(mgmt_rep_recruitment~mgmt_rep_years,xlim=c(1960,2020),typ="l",lwd=3,col="red",ylim=c(0,200))
for(i in 1:n.sims){
lines(sum.rec.area[c(1:43),i]~c(1976:2018),typ="l",lwd=1,col="black")
}
legend("topleft", legend=c("Management EM Recruitment", "Simulated OM Recruitments"),lty=c(1,1),lwd=c(3,1), col=c("red","black"))
```


## Selectivity in OM  
The following figures show selectivity at age for the OPM - these are used for all years of the OM, not just the conditioning period.  
US fixed gear pre IFQ selectivity  
```{r, echo=FALSE}
#selectivty OM vs Mgmt EM
#by fleet for fishery
#US fixed gear pre IFQ
par(mfrow=c(3,2),mar=c(4,4,2,2))
for(m in 1:n.area){
plot(va[1,m,1,]~ages,typ="l",col="black",lwd=3,ylim=c(0,1),ylab="Selectivity",xlab="Age")
lines(va[1,m,2,]~ages,typ="l",col="black",lwd=3,lty=3)
lines(mgmt_rep$Fishery_sel1Females~ages,typ="l",col="red",lwd=3,lty=1)
lines(mgmt_rep$Fishery_sel1Males~ages,typ="l",col="red",lwd=3,lty=3)
title(paste("Area",m,"Fishery",fish[1]))
}
legend("bottomright", legend=c("Management EM Female", "Management EM Male","Simulated OM Female","Simulated OM Male"),lty=c(1,3,1,3),lwd=c(3,3,3,3), col=c("red","red","black","black"),cex=0.75)

```

US fixed gear post IFQ  
```{r, echo=FALSE}  
#US fixed gear post IFQ
par(mfrow=c(3,2),mar=c(4,4,2,2))
for(m in 1:n.area){
plot(va[2,m,1,]~ages,typ="l",col="black",lwd=3,ylim=c(0,1),ylab="Selectivity",xlab="Age")
lines(va[2,m,2,]~ages,typ="l",col="black",lwd=3,lty=3)
lines(mgmt_rep$Fishery_sel4_f~ages,typ="l",col="red",lwd=3,lty=1)
lines(mgmt_rep$Fishery_sel4_m~ages,typ="l",col="red",lwd=3,lty=3)
title(paste("Area",m,"Fishery",fish[2]))
} 
legend("bottomright", legend=c("Management EM Female", "Management EM Male","Simulated OM Female","Simulated OM Male"),lty=c(1,3,1,3),lwd=c(3,3,3,3), col=c("red","red","black","black"),cex=0.75)

```  

US trawl fish  
```{r, echo=FALSE}
#US trawl fish
par(mfrow=c(3,2),mar=c(4,4,2,2))
for(m in 1:n.area){
plot(va[3,m,1,]~ages,typ="l",col="black",lwd=3,ylim=c(0,1),ylab="Selectivity",xlab="Age")
lines(va[3,m,2,]~ages,typ="l",col="black",lwd=3,lty=3)
lines(mgmt_rep$Fishery_sel3~ages,typ="l",col="red",lwd=3,lty=1)
title(paste("Area",m,"Fishery",fish[3]))
} 
legend("topright", legend=c("Management EM Female", "Management EM Male","Simulated OM Female","Simulated OM Male"),lty=c(1,3,1,3),lwd=c(3,3,3,3), col=c("red","red","black","black"),cex=0.75)

```  
  
US fixed gear foreign    
```{r, echo=FALSE}  
#US fixed gear foreign
par(mfrow=c(3,2),mar=c(4,4,2,2))
for(m in 1:n.area){
plot(va[4,m,1,]~ages,typ="l",col="black",lwd=3,ylim=c(0,1),ylab="Selectivity",xlab="Age")
lines(va[4,m,2,]~ages,typ="l",col="black",lwd=3,lty=3)
lines(mgmt_rep$Fishery_sel2~ages,typ="l",col="red",lwd=3,lty=1)
title(paste("Area",m,"Fishery",fish[4]))
}
legend("bottomright", legend=c("Management EM Female", "Management EM Male","Simulated OM Female","Simulated OM Male"),lty=c(1,3,1,3),lwd=c(3,3,3,3), col=c("red","red","black","black"),cex=0.75)

```  

US longline survey - OM selectivity not spatial
```{r, echo=FALSE}  
#US ll survey
par(mfrow=c(1,1),mar=c(4,4,2,2))
plot(va_surv[1,1,]~ages,typ="l",col="black",lwd=3,ylim=c(0,1),ylab="Selectivity",xlab="Age")
lines(va_surv[1,2,]~ages,typ="l",col="black",lwd=3,lty=3)
lines(mgmt_rep$Survey_sel1Female~ages,typ="l",col="red",lwd=3,lty=1)
lines(mgmt_rep$Survey_sel1male~ages,typ="l",col="orange",lwd=3,lty=3)
title(paste("Area","Survey","US LL"))
legend("bottomright", legend=c("Management EM Female", "Management EM Male","Simulated OM Female","Simulated OM Male"),lty=c(1,3,1,3),lwd=c(3,3,3,3), col=c("red","red","black","black"),cex=0.75)

```

USJP ll coop survey  
```{r, echo=FALSE}  
#USJP ll coop survey
par(mfrow=c(1,1),mar=c(4,4,2,2))
plot(va_surv[2,1,]~ages,typ="l",col="black",lwd=3,ylim=c(0,1),ylab="Selectivity",xlab="Age")
lines(va_surv[2,2,]~ages,typ="l",col="black",lwd=3,lty=3)
lines(mgmt_rep$Survey_sel2Female~ages,typ="l",col="red",lwd=3,lty=1)
lines(mgmt_rep$Survey_sel2male~ages,typ="l",col="orange",lwd=3,lty=3)
title(paste("Area","Survey", "USJP coop"))
legend("bottomright", legend=c("Management EM Female", "Management EM Male","Simulated OM Female","Simulated OM Male"),lty=c(1,3,1,3),lwd=c(3,3,3,3), col=c("red","red","black","black"),cex=0.75)

```

## OM sampling for abundance indices, age comps
Need some text here describing the index sampling code (sampling N or B, combining over areas for input to EM, CVs)
   
Need some text here describing the age comp sampling code...  

  
#EM specifications
For 2019 onward, the EM is run at every annual step using data 'sampled' from the OM population. As a reminder the order of operations for year y are: OM - Read in previous year's apportioned ABC by area (from the EM), estimate the F required to catch ABC for each area, apply F and M to OM population, move fish between areas, sample the population for fishery and longline survey abundance indices and for longline survey and fixed gear fishery age compositions, build the .dat file and pass it to the EM in ADMB, run the EM and get an estimate of the next year's ABC (then repeat for the next year).

# Forward looping model output

## EM Convergence
First, we look at the proportion of years within a simulation and apportionment scenario which converged. 
```{r,echo=FALSE}
mgc_val <- 0.01
```

For these simulations, a model was considered 'converged' for a given year if the max gradient component was < `r mgc_val`. The proportion of sims which converged for each year for each apportionment method are:

```{r, echo=FALSE}
AM_max_grads <- array(data=NA, dim=c(n.year,n.sims,length(1:n.apports)), dimnames=list(years,sims,apports)) #Apportionment Master (AM) file for analyses
for(b in 1:n.apports){
  dir.temp <- paste0(dir.output,"/Apport.Option_",b)
AM_max_grads[,,b] <- readRDS(file.path(dir.temp,"maxgrads.rds")) 
}
#proportion of years in a sim that converged
#average proportion of years in all sims that converged for each apportionment option
conv_num <- array(data=NA, dim=c(n.year,n.sims,n.apports), dimnames=list(years,sims,apport.names))
#max_grads <- (format(max_grads,scientific=FALSE))
for(b in 1:n.apports){
  for(i in 1:n.sims){
    for(y in forproj.styr:n.year){
      if(abs(AM_max_grads[y,i,b]) < mgc_val) {
        conv_num[y,i,b] <- 1      
      } else {
        conv_num[y,i,b] <- 0      
      }
    }
  }
}
#proportion converged 
#(proportion of sims that converged in year y) for each apportionment type
temp_a <- apply(conv_num[forproj.styr:n.year,,],c(1,3),sum)

(temp_a1 <- temp_a/n.sims) 
```


The proportion of years which converged for each sims and apportionment method are:
```{r, echo=FALSE}
#(proportion of years that converged in each sim i)
temp_b <- apply(conv_num[forproj.styr:n.year,,],c(2,3),sum)
(temp_b1 <- temp_b/length(forproj.styr:n.year))
```

Finally, the overall proportion of years x sims that converged for each apportionment method are below. Note that for now, all models (converged or not) are included in the following output figures and performance metric calculations.
```{r, echo=FALSE} 
#proportion of sims and years that converged
for(b in 1:n.apports){
temp_c <- sum(temp_b[,b])/(length(forproj.styr:n.year)*n.sims) 
#dimnames(temp_c) <- (apport.names)
print(temp_c)
}
```

## Objective function values
The objective function values for each sim and year iteration of the EM were:  
``` {r, echo=FALSE}
#Apportionment Master (AM) file of objective function values
AM_ofv <- array(data=NA, dim=c(n.year,n.sims,length(1:n.apports)),dimnames=list(c(1976:(1976+n.year-1)),sims,apport.names))  
#Apportionment Master (AM) file of SPR penalty values
AM_sprpen <- array(data=NA, dim=c(n.year,n.sims,length(1:n.apports)), dimnames=list(c(1976:(1976+n.year-1)),sims,apport.names))
#read them in:
for(b in 1:n.apports){
  dir.temp <- paste0(dir.output,"/Apport.Option_",b)
AM_ofv[,,b] <- readRDS(file.path(dir.temp,"obj_fun_vals.rds")) 
AM_sprpen[,,b] <- readRDS(file.path(dir.temp,"spr_penalty.rds")) 
}
AM_ofv[44:n.year,,]
AM_sprpen[44:n.year,,]
```

## EM Selectivity  
``` {r,echo=FALSE}
#create arrays for the selectivities across sims
AM_pred.sel.preifq <- array(data=NA,dim=c(n.year,n.age,n.sex,n.sims,length(1:n.apports)),dimnames=list(c(1976:(1976+n.year-1)),ages,sexes,sims,apport.names))
AM_pred.sel.postifq <- array(data=NA,dim=c(n.year,n.age,n.sex,n.sims,length(1:n.apports)),dimnames=list(c(1976:(1976+n.year-1)),ages,sexes,sims,apport.names))
AM_pred.sel.trawl <- array(data=NA,dim=c(n.year,n.age,n.sex,n.sims,length(1:n.apports)),dimnames=list(c(1976:(1976+n.year-1)),ages,sexes,sims,apport.names))
AM_pred.sel.LLsurv <- array(data=NA,dim=c(n.year,n.age,n.sex,n.sims,length(1:n.apports)),dimnames=list(c(1976:(1976+n.year-1)),ages,sexes,sims,apport.names))
#then read them all in
for(b in 1:n.apports){
  dir.temp <- paste0(dir.output,"/Apport.Option_",b)
AM_pred.sel.preifq[,,,,b] <- readRDS(file.path(dir.temp,"EM_pred.sel.preifqfish.rds")) 
AM_pred.sel.postifq[,,,,b] <- readRDS(file.path(dir.temp,"EM_pred.sel.postifqfish.rds")) 
AM_pred.sel.trawl[,,,,b] <- readRDS(file.path(dir.temp,"EM_pred.sel.trawlfish.rds")) 
AM_pred.sel.LLsurv[,,,,b] <- readRDS(file.path(dir.temp,"EM_pred.sel.LLsurv.rds")) 
}
```

### Longline fishery Pre-IFQ  
EM estimated selectivity for the longline fishery, Pre-IFQ years   
```{r, echo=FALSE}
if(n.apports==1){
par(mfrow=c(1,1))
}
if(n.apports==2){
par(mfrow=c(1,2))
}
if(n.apports>3){
par(mfrow=c(2,2))
}
#females
for(b in 1:n.apports){
plot(va[1,1,1,]~ages,typ="l",lwd=4,ylim=c(0,1),main=c(paste0("Apportionment: ",apport.names[b],", Female LL Pre-IFQ")),ylab="Selectivity",xlab="Age",cex.main=0.75)  #OM selectivity
lines(va[1,2,1,]~ages,typ="l",lwd=4)
lines(va[1,3,1,]~ages,typ="l",lwd=4)
lines(va[1,4,1,]~ages,typ="l",lwd=4)
lines(va[1,5,1,]~ages,typ="l",lwd=4)
lines(va[1,6,1,]~ages,typ="l",lwd=4)
for(y in forproj.styr:n.year){
for(i in 1:n.sims){
lines(AM_pred.sel.preifq[y,,1,i,b]~ages,typ="l",lwd=2, col="red")
}}
legend("bottomright",legend=c("EM","OM"),lwd=c(2,4),col=c("red","black"),cex=0.5)
}
#males
for(b in 1:n.apports){
plot(va[1,1,2,]~ages,typ="l",lwd=4,ylim=c(0,1),main=c(paste0("Apportionment: ",apport.names[b],", Male LL Pre-IFQ")),ylab="Selectivity",xlab="Age",cex.main=0.75)  #OM selectivity
lines(va[1,2,2,]~ages,typ="l",lwd=4)
lines(va[1,3,2,]~ages,typ="l",lwd=4)
lines(va[1,4,2,]~ages,typ="l",lwd=4)
lines(va[1,5,2,]~ages,typ="l",lwd=4)
lines(va[1,6,2,]~ages,typ="l",lwd=4)
for(y in forproj.styr:n.year){
for(i in 1:n.sims){
lines(AM_pred.sel.preifq[y,,2,i,b]~ages,typ="l",lwd=2, col="red")
}}
legend("bottomright",legend=c("EM","OM"),lwd=c(2,4),col=c("red","black"),cex=0.5) 
}

```
  
### Longline fishery Post-IFQ  
```{r, echo=FALSE}
if(n.apports==1){
par(mfrow=c(1,1))
}
if(n.apports==2){
par(mfrow=c(1,2))
}
if(n.apports>3){
par(mfrow=c(2,2))
}
#females
for(b in 1:n.apports){
plot(va[2,1,1,]~ages,typ="l",lwd=4,ylim=c(0,1),main=c(paste0("Apportionment: ",apport.names[b],", Female LL Post-IFQ")),ylab="Selectivity",xlab="Age",cex.main=0.75)  #OM selectivity
lines(va[2,2,1,]~ages,typ="l",lwd=4)
lines(va[2,3,1,]~ages,typ="l",lwd=4)
lines(va[2,4,1,]~ages,typ="l",lwd=4)
lines(va[2,5,1,]~ages,typ="l",lwd=4)
lines(va[2,6,1,]~ages,typ="l",lwd=4)
for(y in forproj.styr:n.year){
for(i in 1:n.sims){
lines(AM_pred.sel.postifq[y,,1,i,b]~ages,typ="l",lwd=2, col="red")
}}
legend("bottomright",legend=c("EM","OM"),lwd=c(2,4),col=c("red","black"),cex=0.5)
}
#males
for(b in 1:n.apports){
plot(va[2,1,2,]~ages,typ="l",lwd=4,ylim=c(0,1),main=c(paste0("Apportionment: ",apport.names[b],", Male LL Post-IFQ")),ylab="Selectivity",xlab="Age",cex.main=0.75)  #OM selectivity
lines(va[2,2,2,]~ages,typ="l",lwd=4)
lines(va[2,3,2,]~ages,typ="l",lwd=4)
lines(va[2,4,2,]~ages,typ="l",lwd=4)
lines(va[2,5,2,]~ages,typ="l",lwd=4)
lines(va[2,6,2,]~ages,typ="l",lwd=4)
for(y in forproj.styr:n.year){
for(i in 1:n.sims){
lines(AM_pred.sel.postifq[y,,2,i,b]~ages,typ="l",lwd=2, col="red")
}}
legend("bottomright",legend=c("EM","OM"),lwd=c(2,4),col=c("red","black"),cex=0.5)
}
```
  
### Trawl fishery
```{r, echo=FALSE}
if(n.apports==1){
par(mfrow=c(1,1))
}
if(n.apports==2){
par(mfrow=c(1,2))
}
if(n.apports>3){
par(mfrow=c(2,2))
}
#females
for(b in 1:n.apports){
plot(va[3,1,1,]~ages,typ="l",lwd=4,ylim=c(0,1),main=c(paste0("Apportionment: ",apport.names[b],", Female LL Trawl")),ylab="Selectivity",xlab="Age",cex.main=0.75)  #OM selectivity
lines(va[3,2,1,]~ages,typ="l",lwd=4)
lines(va[3,3,1,]~ages,typ="l",lwd=4)
lines(va[3,4,1,]~ages,typ="l",lwd=4)
lines(va[3,5,1,]~ages,typ="l",lwd=4)
lines(va[3,6,1,]~ages,typ="l",lwd=4)
for(y in forproj.styr:n.year){
for(i in 1:n.sims){
lines(AM_pred.sel.trawl[y,,1,i,b]~ages,typ="l",lwd=2, col="red")
}}
legend("topright",legend=c("EM","OM"),lwd=c(2,4),col=c("red","black"),cex=0.5)
}
#males
for(b in 1:n.apports){
plot(va[3,1,2,]~ages,typ="l",lwd=4,ylim=c(0,1),main=c(paste0("Apportionment: ",apport.names[b],", Male LL Trawl")),ylab="Selectivity",xlab="Age",cex.main=0.75)  #OM selectivity
lines(va[3,2,2,]~ages,typ="l",lwd=4)
lines(va[3,3,2,]~ages,typ="l",lwd=4)
lines(va[3,4,2,]~ages,typ="l",lwd=4)
lines(va[3,5,2,]~ages,typ="l",lwd=4)
lines(va[3,6,2,]~ages,typ="l",lwd=4)
for(y in forproj.styr:n.year){
for(i in 1:n.sims){
lines(AM_pred.sel.trawl[y,,2,i,b]~ages,typ="l",lwd=2, col="red")
}}
legend("topright",legend=c("EM","OM"),lwd=c(2,4),col=c("red","black"),cex=0.5)
}
```
  
### Longline survey (US years)
```{r, echo=FALSE}
if(n.apports==1){
par(mfrow=c(1,1))
}
if(n.apports==2){
par(mfrow=c(1,2))
}
if(n.apports>3){
par(mfrow=c(2,2))
}
#females
for(b in 1:n.apports){
plot(va_surv[1,1,]~ages,typ="l",lwd=4,ylim=c(0,1),main=c(paste0("Apportionment: ",apport.names[b],", Female LL Survey")),ylab="Selectivity",xlab="Age",cex.main=0.75)  #OM selectivity
for(y in forproj.styr:n.year){
for(i in 1:n.sims){
lines(AM_pred.sel.LLsurv[y,,1,i,b]~ages,typ="l",lwd=2, col="red")
}}
legend("bottomright",legend=c("EM","OM"),lwd=c(2,4),col=c("red","black"),cex=0.5)
}
#males
for(b in 1:n.apports){
plot(va_surv[1,2,]~ages,typ="l",lwd=4,ylim=c(0,1),main=c(paste0("Apportionment: ",apport.names[b],", Male LL Survey")),ylab="Selectivity",xlab="Age",cex.main=0.75)  #OM selectivity
for(y in forproj.styr:n.year){
for(i in 1:n.sims){
lines(AM_pred.sel.LLsurv[y,,2,i,b]~ages,typ="l",lwd=2, col="red")
}}
legend("bottomright",legend=c("EM","OM"),lwd=c(2,4),col=c("red","black"),cex=0.5)
}
```
  
## Catchability  
Catchability for the EM is shown below in boxplots. For each apportionment option (x axis) the box shows the median (thick line inside the box) and the 25th and 75th percentile interquartile range (box lower and upper border) of EM q estimates across all years and sims. The vertical bars represent the largest and smallest values within 1.5 times the interquartile range, and any values outside these ranges are shown as points. The dashed red line is the OM value of catchability.  
```{r, echo=FALSE}
#create matrices to hold q values from EM for all apportionments
AM_EM_q.LLsurv <- array(data=NA,dim=c(n.year,n.sims,length(1:n.apports)),dimnames=list(c(1976:(1976+n.year-1)),sims,apport.names))
AM_EM_q.USJPsurv <- array(data=NA,dim=c(n.year,n.sims,length(1:n.apports)),dimnames=list(c(1976:(1976+n.year-1)),sims,apport.names))
AM_EM_q.preifqfish <- array(data=NA,dim=c(n.year,n.sims,length(1:n.apports)),dimnames=list(c(1976:(1976+n.year-1)),sims,apport.names))
AM_EM_q.postifqfish <- array(data=NA,dim=c(n.year,n.sims,length(1:n.apports)),dimnames=list(c(1976:(1976+n.year-1)),sims,apport.names))
AM_EM_q.forfish <- array(data=NA,dim=c(n.year,n.sims,length(1:n.apports)),dimnames=list(c(1976:(1976+n.year-1)),sims,apport.names))
OM_q_fish <- array(data=NA,dim=c(n.fish,n.area,n.apports),dimnames=list(fish,areas,apport.names))
OM_q_surv <- array(data=NA,dim=c(n.surv,n.area,n.apports),dimnames=list(surv.name,areas,apport.names))

#then read them all in
for(b in 1:n.apports){
dir.temp <- paste0(dir.output,"/Apport.Option_",b)
AM_EM_q.LLsurv[,,b] <- readRDS(file.path(dir.temp,"EM_q.LLsurv.rds")) 
AM_EM_q.USJPsurv[,,b] <- readRDS(file.path(dir.temp,"EM_q.USJPsurv.rds")) 
AM_EM_q.preifqfish[,,b] <- readRDS(file.path(dir.temp,"EM_q.preifqfish.rds")) 
AM_EM_q.postifqfish[,,b] <- readRDS(file.path(dir.temp,"EM_q.postifqfish.rds")) 
AM_EM_q.forfish[,,b] <- readRDS(file.path(dir.temp,"EM_q.forfish.rds"))
OM_q_fish[,,b] <- readRDS(file.path(dir.temp,"q_fish.rds"))
OM_q_surv[,,b] <- readRDS(file.path(dir.temp,"q_surv.rds"))
}

```
  
### Fixed gear fishery, foreign years  
```{r,echo=FALSE}
melted_EMq.forfish <- melt(AM_EM_q.forfish[forproj.styr:n.year,,],na.rm=FALSE,value.name="Catchability")
ggplot(melted_EMq.forfish,aes(x=Var3,y=Catchability)) +
  geom_boxplot() +
  geom_hline(yintercept=OM_q_fish[4,1,1],linetype="solid",color="red",size=2) 
```
   
### Longline fishery Pre-IFQ  
For the boxplots below, the dashed red line is OM q for BS, AI, and WG. The solid red line is OM q for CG, and the dotted line is OM q for WY and EY/SEO.   
  
```{r,echo=FALSE}
melted_EMq.preifq <- melt(AM_EM_q.preifqfish[forproj.styr:n.year,,],na.rm=FALSE,value.name="Catchability")
ggplot(melted_EMq.preifq,aes(x=Var3,y=Catchability)) +
  geom_boxplot() +
  geom_hline(yintercept=OM_q_fish[1,1,1],linetype="dashed",color="red",size=2) +
  geom_hline(yintercept=OM_q_fish[1,2,1],linetype="dashed",color="red",size=2) +
  geom_hline(yintercept=OM_q_fish[1,3,1],linetype="dashed",color="red",size=2) + 
  geom_hline(yintercept=OM_q_fish[1,4,1],linetype="solid",color="red",size=2) + 
  geom_hline(yintercept=OM_q_fish[1,5,1],linetype="dotted",color="red",size=2) +
  geom_hline(yintercept=OM_q_fish[1,6,1],linetype="dotted",color="red",size=2) 
```
   
### Longline fishery Post-IFQ  
For the boxplots below, the dashed red line is OM q for BS, AI, and WG. The solid red line is OM q for CG, and the dotted line is OM q for WY and EY/SEO.    
```{r,echo=FALSE}
melted_EMq.postifq <- melt(AM_EM_q.postifqfish[forproj.styr:n.year,,],na.rm=FALSE,value.name="Catchability")
ggplot(melted_EMq.postifq,aes(x=Var3,y=Catchability)) +
  geom_boxplot() +
  geom_hline(yintercept=OM_q_fish[2,1,1],linetype="dashed",color="red",size=2) +
  geom_hline(yintercept=OM_q_fish[2,2,1],linetype="dashed",color="red",size=2) +
  geom_hline(yintercept=OM_q_fish[2,3,1],linetype="dashed",color="red",size=2) + 
  geom_hline(yintercept=OM_q_fish[2,4,1],linetype="solid",color="red",size=2) + 
  geom_hline(yintercept=OM_q_fish[2,5,1],linetype="dotted",color="red",size=2) +
  geom_hline(yintercept=OM_q_fish[2,6,1],linetype="dotted",color="red",size=2) 
```
   
### Longline survey (US years)     
```{r,echo=FALSE}
melted_EMq.LLsurv <- melt(AM_EM_q.LLsurv[forproj.styr:n.year,,],na.rm=FALSE,value.name="Catchability")
ggplot(melted_EMq.LLsurv,aes(x=Var3,y=Catchability)) +
  geom_boxplot() +
  geom_hline(yintercept=OM_q_surv[1,1,1],linetype="dashed",color="red",size=3)

```
    
### Longline survey (USJP years)
```{r,echo=FALSE}
melted_EMq.USJPsurv <- melt(AM_EM_q.USJPsurv[forproj.styr:n.year,,],na.rm=FALSE,value.name="Catchability")
ggplot(melted_EMq.USJPsurv,aes(x=Var3,y=Catchability)) +
  geom_boxplot() +
  geom_hline(yintercept=OM_q_surv[2,1,1],linetype="dashed",color="red",size=3)

```
   

## Indices  
Do we need to add a figure or analysis showing the EM fit to OM true population, beyond just the fit to the OM sampled indices?
``` {r, echo=FALSE}
#create arrays to hold indices
AM_EM_pred.srvRPN <- array(data=NA,dim=c(n.year,n.year,n.sims,length(1:n.apports)),dimnames=list(years,years,sims,apport.names))
AM_EM_pred.fishRPW <- array(data=NA,dim=c(n.year,n.year,n.sims,length(1:n.apports)),dimnames=list(years,years,sims,apport.names))
OM.Surv.RPN <- array(data=NA,dim=c(n.year,n.sims,length(1:n.apports)),dimnames=list(years,sims,apport.names))
OM.Fish.RPW <- array(data=NA,dim=c(n.year,n.sims,length(1:n.apports)),dimnames=list(years,sims,apport.names))
#read in the index predictions from the EM and OM
for(b in 1:n.apports){
dir.temp <- paste0(dir.output,"/Apport.Option_",b)
AM_EM_pred.srvRPN[,,,b] <- readRDS(file.path(dir.temp,"EM_pred.srvRPN.rds")) 
AM_EM_pred.fishRPW[,,,b] <- readRDS(file.path(dir.temp,"EM_pred.fishRPW.rds")) 
OM.Surv.RPN[,,b] <- readRDS(file.path(dir.temp,"OM_Surv.RPN.rds")) 
OM.Fish.RPW[,,b] <- readRDS(file.path(dir.temp,"OM_Fish.RPW.rds")) 
}
```

### US longline survey index (RPN)   
The OM ('true') and EM fits to the longline survey abundance index are shown below for each apportionment method. 
``` {r, echo=FALSE}
#melted_EM_pred.srvRPN <- melt(AM_EM_pred.srvRPN,na.rm=TRUE,value.name="EM.RPN") 
#Plot EM and OM RPN terminal year only, across sims
if(n.apports==1){
par(mfrow=c(1,1))
}
if(n.apports==2){
par(mfrow=c(1,2))
}
if(n.apports>3){
par(mfrow=c(2,2))
}
for(b in 1:n.apports){
plot(AM_EM_pred.srvRPN[n.year,15:n.year,1,b]~c(15:n.year),typ="l",ylim=c(0,2000),ylab="Survey RPN",xlab="Year",col="red",main=apport.names[b])
#for(y in forproj.styr:n.year){
  for(i in 1:n.sims){
    lines(AM_EM_pred.srvRPN[n.year,15:n.year,i,b]~c(15:n.year),typ="l",col="red")
  }
for(i in 1:n.sims){
  lines(OM.Surv.RPN[15:n.year,i,b]~c(15:n.year),typ="l",col="black")
}
  legend("topleft",legend=c("EM","OM"),lwd=c(2,4),col=c("red","black"),cex=0.5)
}

```
  

``` {r, echo=FALSE}
#OM-EM Residuals - these are the residuals between estimated EM and 'true' EM survey index for each apportionment method, using all years and sims (not just the terminal year EM run index estimates)
#resid.RPN <- array(data=NA, dim=c(n.year,n.year,n.sims,n.apports),dimnames=list(years,years,sims,apport.names))
#for(b in 1:n.apports) {
#for (y in 1:n.year) {
#for(k in 1:n.year) {
# for(i in 1:n.sims) {
#resid.RPN[y,k,i,b] <- OM.Surv.RPN[y,i,b] - AM_EM_pred.srvRPN[y,k,i,b]
#}}}}
#melt_resid.rpn <- melt(resid.RPN,na.rm=TRUE,value.name="resid")
#ggplot(melt_resid.rpn,aes(x=Var4,y=resid))+
#geom_boxplot()

```
  
Just using the terminal year EM run residuals from OM, for all sims.  Do these make sense??  
``` {r, echo=FALSE}
resid.RPN2 <- array(data=NA, dim=c(n.year,n.sims,n.apports),dimnames=list(years,sims,apport.names))
for(b in 1:n.apports) {
#for (y in 1:n.year) {
  for(k in 1:n.year) {
  for(i in 1:n.sims) {
    resid.RPN2[k,i,b] <- OM.Surv.RPN[k,i,b] - AM_EM_pred.srvRPN[n.year,k,i,b]
  }}}#}
#resid.RPN[forproj.styr:n.year,15:n.year,,]
melt_resid.rpn2 <- melt(resid.RPN2,na.rm=TRUE,value.name="resid")
ggplot(melt_resid.rpn2,aes(x=Var3,y=resid))+
  geom_boxplot()

```
  
### US longline fishery index (RPW)  
The OM ('true') and EM fits to the longline fishery abundance index are shown below for each apportionment method. Why doesn't the OM want to get as high(for samples) as the EM? Recruitment??
``` {r, echo=FALSE}
#melted_EM_pred.srvRPN <- melt(AM_EM_pred.srvRPN,na.rm=TRUE,value.name="EM.RPN") 
#Plot EM and OM RPN terminal year only, across sims
if(n.apports==1){
par(mfrow=c(1,1))
}
if(n.apports==2){
par(mfrow=c(1,2))
}
if(n.apports>3){
par(mfrow=c(2,2))
}
for(b in 1:n.apports){
plot(AM_EM_pred.fishRPW[n.year,15:n.year,1,b]~c(15:n.year),typ="l",ylim=c(500,2200),ylab="Fishery RPW",xlab="Year",col="red",main=apport.names[b])
#for(y in forproj.styr:n.year){
  for(i in 1:n.sims){
    lines(AM_EM_pred.fishRPW[n.year,15:n.year,i,b]~c(15:n.year),typ="l",col="red")
  }
for(i in 1:n.sims){
  lines(OM.Fish.RPW[15:n.year,i,b]~c(15:n.year),typ="l",col="black")
}
  legend("topleft",legend=c("EM","OM"),lwd=c(2,4),col=c("red","black"),cex=0.5)
}

```
  
Using the terminal year EM run residuals from OM, for all sims.  Do these make sense??  
``` {r, echo=FALSE}
resid.RPW2 <- array(data=NA, dim=c(n.year,n.sims,n.apports),dimnames=list(years,sims,apport.names))
for(b in 1:n.apports) {
#for (y in 1:n.year) {
  for(k in 1:n.year) {
  for(i in 1:n.sims) {
    resid.RPW2[k,i,b] <- OM.Fish.RPW[k,i,b] - AM_EM_pred.fishRPW[n.year,k,i,b]
  }}}#}
#resid.RPN[forproj.styr:n.year,15:n.year,,]
melt_resid.rpw2 <- melt(resid.RPW2,na.rm=TRUE,value.name="resid")
ggplot(melt_resid.rpw2,aes(x=Var3,y=resid))+
  geom_boxplot()

```
  
## Recruitment  
``` {r,echo=FALSE}
#create arrays to hold indices
AM_EM_predrec <- array(data=NA,dim=c(n.year,n.year,n.sims,length(1:n.apports)),dimnames=list(years,years,sims,apport.names))
OM.predrec <- array(data=NA,dim=c(n.year,n.area,n.sims,length(1:n.apports)),dimnames=list(years,areas,sims,apport.names))
OM_N <- array(data=NA,dim=c(n.sex,n.year,n.age,n.area,n.sims,length(1:n.apports)),dimnames=list(sexes,years,ages,areas,sims,apport.names))
#read in the index predictions from the EM and OM
for(b in 1:n.apports){
dir.temp <- paste0(dir.output,"/Apport.Option_",b)
AM_EM_predrec[,,,b] <- readRDS(file.path(dir.temp,"EM_predrec.rds")) 
OM.predrec[,,,b] <- readRDS(file.path(dir.temp,"recruits.area.rds")) 
OM_N[,,,,,b] <- readRDS(file.path(dir.temp,"N.rds"))
}
```
  
First look at all OM and EM recruitment draws by apport scenario
``` {r, echo=FALSE}
sum_OM.predrec <- apply(OM.predrec,c(1,3,4),sum) #sum OM predicted recruitment across areas
if(n.apports==1){
par(mfrow=c(1,1))
}
if(n.apports==2){
par(mfrow=c(1,2))
}
if(n.apports>3){
par(mfrow=c(2,2))
}
for(b in 1:n.apports){
plot(AM_EM_predrec[n.year,,1,b]~years,col="red",lwd=1,typ="l", ylab="Predicted Recruitment (millions)",xlab="Year",main=apport.names[b])
  for(i in 1:n.sims){
  lines(AM_EM_predrec[n.year,,i,b]~years,col="red",lwd=1,typ="l")
    lines(sum_OM.predrec[,i,b]~years,col="black",lwd=1,typ="l")
  }
  legend("topleft",legend=c("EM","OM"),lwd=c(1,1),col=c("red","black"),cex=0.5)
  }
```
  
Then look at the median EM and median OM recruitment draws by apport scenario. What does this mean that these are so different?  
``` {r, echo=FALSE}
EM_recmed <- apply(AM_EM_predrec[n.year,,,],c(1,3),median)
OM_recmed <- apply(sum_OM.predrec,c(1,3),median)
if(n.apports==1){
par(mfrow=c(1,1))
}
if(n.apports==2){
par(mfrow=c(1,2))
}
if(n.apports>3){
par(mfrow=c(2,2))
}
for(b in 1:n.apports){
plot(EM_recmed[,b]~years,col="red",lwd=1,typ="l",main=paste(apport.names[b],"- Median"),ylab="Recruitment (millions)",xlab="Year")
lines(OM_recmed[,b]~years,col="black",lwd=1,typ="l")
legend("topleft",legend=c("EM","OM"),lwd=c(1,1),col=c("red","black"),cex=0.5)
}
```

Then look at the mean EM and mean OM recruitment draws by apport scenario. These are a bit more similar.
``` {r, echo=FALSE}
if(n.apports==1){
par(mfrow=c(1,1))
}
if(n.apports==2){
par(mfrow=c(1,2))
}
if(n.apports>3){
par(mfrow=c(2,2))
}
EM_recmean <- apply(AM_EM_predrec[n.year,,,],c(1,3),mean)
OM_recmean <- apply(sum_OM.predrec,c(1,3),mean)
for(b in 1:n.apports){
plot(EM_recmean[,b]~years,col="red",lwd=1,typ="l",main=paste(apport.names[b],"- Mean"),ylab="Recruitment (millions)",xlab="Year")
lines(OM_recmean[,b]~years,col="black",lwd=1,typ="l")
legend("topleft",legend=c("EM","OM"),lwd=c(1,1),col=c("red","black"),cex=0.5)
}
```
  
Then look at mean EM recruiment, mean OM recruitment, and OM N at age 2.
Why is the EM estimating such high recruitments in the projection years (years 44+) when it doesn't seem to be coming from the OM?  What would be driving that?  Look at residuals next.
``` {r, echo=FALSE}
sum_OM_N <- apply(OM_N[,,1,,,],c(2,4,5),sum) #N at age 2, summed over areas
if(n.apports==1){
par(mfrow=c(1,1))
}
if(n.apports==2){
par(mfrow=c(1,2))
}
if(n.apports>3){
par(mfrow=c(2,2))
}
#par(mfrow=c(1,1))
i=1
for(b in 1:n.apports){
plot(AM_EM_predrec[n.year,,1,b]~years,col="red",lwd=1,typ="l", ylab="Predicted Recruitment (millions)",xlab="Year",main=apport.names[b])
  for(i in 1:n.sims){
  lines(AM_EM_predrec[n.year,,i,b]~years,col="red",lwd=3,typ="l")
  lines(sum_OM.predrec[forproj.styr:n.year,i,b]~c(forproj.styr:n.year),col="black",lwd=5,lty=1)
  lines(sum_OM_N[forproj.styr:n.year,i,b]~c(forproj.styr:n.year), col="green",lwd=3,lty=3)
  }
  legend("topleft",legend=c("EM pred rec","OM rec","OM N at age 2"),lwd=c(3,3,3),col=c("red","black","green"),cex=0.5)
  }
```

Then look at EM-OM residuals  
``` {r, echo=FALSE}
resid.rec2 <- array(data=NA, dim=c(n.year,n.sims,n.apports),dimnames=list(years,sims,apport.names))
for(b in 1:n.apports) {
#for (y in 1:n.year) {
  for(k in 1:n.year) {
  for(i in 1:n.sims) {
    resid.rec2[k,i,b] <- sum_OM_N[k,i,b] - AM_EM_predrec[n.year,k,i,b]
  }}}#}
#resid.RPN[forproj.styr:n.year,15:n.year,,]
melt_resid.rec2 <- melt(resid.rec2,na.rm=TRUE,value.name="resid")
ggplot(melt_resid.rec2,aes(x=Var3,y=resid))+
  geom_boxplot()
```
  
What are other ways of looking at recruitment that might be interesting?
  

## SSB time series  
``` {r, echo=FALSE}
#create arrays for the selectivities across sims
AM_EM_spbiom <- array(data=NA,dim=c(n.year,n.year,n.sims,n.apports),dimnames=list(years,years,sims,apport.names))
OM_ssb <- array(data=NA,dim=c(n.sex,n.age,n.year,n.area,n.sims,length(1:n.apports)),dimnames=list(sexes,ages,years,areas,sims,apport.names))
#then read them all in
for(b in 1:n.apports){
  dir.temp <- paste0(dir.output,"/Apport.Option_",b)
AM_EM_spbiom[,,,b] <- readRDS(file.path(dir.temp,"EM_spbiom.rds")) 
OM_ssb[,,,,,b] <- readRDS(file.path(dir.temp,"ssb.rds")) 
}
```
  
Plot OM spawning biomass vs EM spawning biomass (for all the terminal sim years and sims), for each apportionment method  
``` {r, echo=FALSE}
sum_OM_ssb <- apply(OM_ssb,c(3,5,6),sum)
if(n.apports==1){
par(mfrow=c(1,1))
}
if(n.apports==2){
par(mfrow=c(1,2))
}
if(n.apports>3){
par(mfrow=c(2,2))
}
for(b in 1:n.apports){
plot(sum_OM_ssb[,1,b]~years,col="black",lwd=1,typ="l",ylim=c(0,300),ylab=c("SSB(kt)"),xlab=c("Year"),main=apport.names[b])
for(i in 1:n.sims){
lines(AM_EM_spbiom[n.year,,i,b]~years,col="red",lwd=1,typ="l")
legend("topright",legend=c("EM SSB","OM SSB"),lwd=c(3,3),col=c("red","black"),cex=0.5)
}}
```

SSB residuals 
Two ways.  
```{r,echo=FALSE}
resid.ssb2 <- array(data=NA, dim=c(n.year,n.sims,n.apports),dimnames=list(years,sims,apport.names))
for(b in 1:n.apports) {
#for (y in 1:n.year) {
  for(k in 1:n.year) {
  for(i in 1:n.sims) {
    resid.ssb2[k,i,b] <- sum_OM_ssb[k,i,b] - AM_EM_spbiom[n.year,k,i,b]
  }}}#}
#resid.RPN[forproj.styr:n.year,15:n.year,,]
melt_resid.ssb2 <- melt(resid.ssb2,na.rm=TRUE,value.name="resid")
names(melt_resid.ssb2) <- c("Year","Sim","Apportionment","Resid")
ggplot(melt_resid.ssb2,aes(x=Apportionment,y=Resid))+
  geom_boxplot()

melt_resid.ssb2 %>% 
  ggplot(aes(x=factor(Year),y=Resid,fill=Apportionment)) + 
           geom_boxplot()

```

  
# Performance metrics   
```{r,echo=FALSE}
#create arrays for the apportionments
AM_EM_abc <- array(data=NA,dim=c(n.year,n.fish,n.area,n.sims,n.apports),dimnames=list(years,fish,areas,sims,apport.names))
#then read them all in
for(b in 1:n.apports){
  dir.temp <- paste0(dir.output,"/Apport.Option_",b)
AM_EM_abc[,,,,b] <- readRDS(file.path(dir.temp,"apportioned_C.rds")) 
}
```

The overall mean and median (areas summed and years summed, for each apportionment method)  
``` {r,echo=FALSE}
abc_allmean <- apply(AM_EM_abc,5,mean)
abc_allmed <- apply(AM_EM_abc,5,median)
abc_allmean
abc_allmed

allmean_sd <- vector()
allmean_uci <- vector()
allmean_lci <- vector()
for(b in 1:n.apports){
allmean_sd[b] <- sd(AM_EM_abc[,,,,b])
allmean_uci[b] <- abc_allmean[b] + (1.92*allmean_sd[b])
allmean_lci[b] <- abc_allmean[b] - (1.92*allmean_sd[b])
}

#make a matrix of means, medians, uci, lci
all_abc <- matrix(data=NA,nrow=n.apports,ncol=4)
colnames(all_abc) <- c("Mean","Median","UCI","LCI")
rownames(all_abc) <- c(apport.names)
for(b in 1:n.apports){
all_abc[b,1] <- abc_allmean[b]
all_abc[b,2] <- abc_allmed[b]
all_abc[b,3] <- allmean_uci[b]
all_abc[b,4] <- allmean_lci[b]
}

melted_ABC <- melt(AM_EM_abc[43:n.year,,,,], value.name="ABC")
names(melted_ABC) <- c("Year","Fishery","Area","Sim","Apportionment.Method","ABC")
melted_ABC %>% 
ggplot(aes(x=factor(Apportionment.Method),y=ABC)) +#,fill=Apportionment)) + 
           geom_boxplot()

```
  
  
Median and mean apportioned ABC by area (over all sims and forward looping years) - relative means and medians
```{r,echo=FALSE}
#set up objects to hold secondary products
mean_ABC <- array(data=NA,dim=c(n.year,n.area,n.apports),dimnames=list(years,areas,apport.names))
median_ABC <- array(data=NA,dim=c(n.year,n.area,n.apports),dimnames=list(years,areas,apport.names))
meanABC_uci <- array(data=NA,dim=c(n.year,n.area,n.apports),dimnames=list(years,areas,apport.names))
meanABC_lci <- array(data=NA,dim=c(n.year,n.area,n.apports),dimnames=list(years,areas,apport.names))
#medianABC_uci <- array(data=NA,dim=c(n.year,n.area,n.apports),dimnames=list(years,areas,apport.names))
#medianABC_lci <- array(data=NA,dim=c(n.year,n.area,n.apports),dimnames=list(years,areas,apport.names))
sd_1 <- array(data=NA,dim=c(n.year,n.area,n.apports),dimnames=list(years,areas,apport.names))
#calculate things
mean_ABC <- apply(AM_EM_abc,c(1,3,5),mean)
median_ABC <- apply(AM_EM_abc,c(1,3,5),median)

#calc sd for year, area, apportionment, across sims
#first, sum abc over gear and take sd
for(y in 1:n.year){
  for(m in 1:n.area) {
    for(b in 1:n.apports){
sd_1[y,m,b] <- sd((apply(AM_EM_abc,c(1,3,4,5),sum))[y,m,,b])
    }}}
#then fill in uci and lci 
for(y in 1:n.year){
  for(m in 1:n.area) {
    for(b in 1:n.apports){
meanABC_lci[y,m,b] <- mean_ABC[y,m,b] - (1.96*sd_1[y,m,b])
meanABC_uci[y,m,b] <- mean_ABC[y,m,b] + (1.96*sd_1[y,m,b])
    }}}  
#then plot
par(mfrow=c(3,2))
for(b in 1:n.apports){
  for(m in 1:n.area){
plot(mean_ABC[(forproj.styr-1):n.year,m,b]~years[(forproj.styr-1):n.year],typ="l",col="black",lwd=3,ylim=c(0,6),ylab=c("ABC (kt)"), xlab=c("Sim year"),main=paste(c(apport.names[b], "Area", areas[m])))
lines(median_ABC[(forproj.styr-1):n.year,m,b]~years[(forproj.styr-1):n.year],typ="l",col="red",lwd=3)
lines(meanABC_uci[(forproj.styr-1):n.year,m,b]~years[(forproj.styr-1):n.year],lty=3,col="grey",lwd=1)
lines(meanABC_lci[(forproj.styr-1):n.year,m,b]~years[(forproj.styr-1):n.year],lty=3,col="grey",lwd=1)
legend("topright",legend=c("Mean ABC","Median ABC","Upper & Lower ~95% CI"),lwd=c(3,3),col=c("black","red","grey","grey"),cex=0.5)
}}
``` 

Median and mean apportioned ABC by area and by gear (over all sims and forward looping years) - relative means and medians
``` {r,echo=FALSE}
#set up objects to hold secondary products
mean_ABC_gear <- array(data=NA,dim=c(n.year,n.fish,n.area,n.apports),dimnames=list(years,fish, areas,apport.names))
median_ABC_gear <- array(data=NA,dim=c(n.year,n.fish,n.area,n.apports),dimnames=list(years,fish,areas,apport.names))
meanABC_uci_gear <- array(data=NA,dim=c(n.year,n.fish,n.area,n.apports),dimnames=list(years,fish,areas,apport.names))
meanABC_lci_gear <- array(data=NA,dim=c(n.year,n.fish,n.area,n.apports),dimnames=list(years,fish,areas,apport.names))
medianABC_uci_gear <- array(data=NA,dim=c(n.year,n.fish,n.area,n.apports),dimnames=list(years,fish,areas,apport.names))
medianABC_lci_gear <- array(data=NA,dim=c(n.year,n.fish,n.area,n.apports),dimnames=list(years,fish,areas,apport.names))
sd_2 <- array(data=NA,dim=c(n.year,n.fish,n.area,n.apports),dimnames=list(years,fish,areas,apport.names))
#calculate things
mean_ABC_gear <- apply(AM_EM_abc,c(1,2,3,5),mean)
median_ABC_gear <- apply(AM_EM_abc,c(1,2,3,5),median)

#calc sd for year, area, apportionment, across sims
#first, sum abc over gear and take sd
for(y in 1:n.year){
  for(f in 1:n.fish){
    for(m in 1:n.area) {
      for(b in 1:n.apports){
sd_2[y,f,m,b] <- sd(AM_EM_abc[y,f,m,,b])
      }}}}
#then fill in uci and lci 
for(y in 1:n.year){
  for(f in 1:n.fish){
    for(m in 1:n.area) {
      for(b in 1:n.apports){
meanABC_lci_gear[y,f,m,b] <- mean_ABC_gear[y,f,m,b] - (1.96*sd_2[y,f,m,b])
meanABC_uci_gear[y,f,m,b] <- mean_ABC_gear[y,f,m,b] + (1.96*sd_2[y,f,m,b])
      }}}} 
#then plot
par(mfrow=c(3,2))
for(b in 1:n.apports){
  for(m in 1:n.area){
plot(mean_ABC_gear[(forproj.styr-1):n.year,2,m,b]~years[(forproj.styr-1):n.year],typ="l",col="black",lwd=3,ylim=c(0,6),ylab=c("ABC (kt)"), xlab=c("Sim year"),main=paste(c(apport.names[b], "Area", areas[m])))
  lines(mean_ABC_gear[(forproj.styr-1):n.year,3,m,b]~years[(forproj.styr-1):n.year],typ="l",col="red",lwd=3)
  lines(median_ABC_gear[(forproj.styr-1):n.year,2,m,b]~years[(forproj.styr-1):n.year],typ="l",col="red",lwd=3)
lines(meanABC_uci_gear[(forproj.styr-1):n.year,m,b]~years[(forproj.styr-1):n.year],lty=3,col="grey",lwd=1)
lines(meanABC_lci_gear[(forproj.styr-1):n.year,m,b]~years[(forproj.styr-1):n.year],lty=3,col="grey",lwd=1)
legend("topright",legend=c("Mean ABC","Median ABC","Upper & Lower ~95% CI"),lwd=c(3,3),col=c("black","red","grey","grey"),cex=0.5)
}}
```

Median and mean apportioned ABC for all areas and gears summed (over all areas, sims and forward looping years) - relative values
``` {r,echo=FALSE}
mean_ABC_allA <- matrix(data=NA,nrow=c(n.year),ncol=c(n.apports),dimnames=list(years,apport.names))
median_ABC_allA <- matrix(data=NA,nrow=c(n.year),ncol=c(n.apports),dimnames=list(years,apport.names))
meanABC_uci_allA <- matrix(data=NA,nrow=c(n.year),ncol=c(n.apports),dimnames=list(years,apport.names))
meanABC_lci_allA <- matrix(data=NA,nrow=c(n.year),ncol=c(n.apports),dimnames=list(years,apport.names))
medianABC_uci_allA <- matrix(data=NA,nrow=c(n.year),ncol=c(n.apports),dimnames=list(years,apport.names))
medianABC_lci_allA <- matrix(data=NA,nrow=c(n.year),ncol=c(n.apports),dimnames=list(years,apport.names))

```

Median and mean apportioned ABC by gear for all areas summed (over all areas, sims and forward looping years) - relative values
``` {r,echo=FALSE}
mean_ABC_allgear <- array(data=NA,dim=c(n.year,n.fish,n.apports),dimnames=list(years,fish,apport.names))
median_ABC_allgear <- array(data=NA,dim=c(n.year,n.fish,n.apports),dimnames=list(years,fish,apport.names))
meanABC_uci_allgear <- array(data=NA,dim=c(n.year,n.fish,n.apports),dimnames=list(years,fish,apport.names))
meanABC_lci_allgear <- array(data=NA,dim=c(n.year,n.fish,n.apports),dimnames=list(years,fish,apport.names))
medianABC_uci_allgear <- array(data=NA,dim=c(n.year,n.fish,n.apports),dimnames=list(years,fish,apport.names))
medianABC_lci_allgear <- array(data=NA,dim=c(n.year,n.fish,n.apports),dimnames=list(years,fish,apport.names))

```

proportion of years (terminal year of EM runs) and sims where apportioned TAC falls below threshold
proportion years in each area above threshold across all years and sims 


## stability for each apportionment option (over all areas summed, and by area) - mean and median absolute change in in ABC year to year and for all years


## How well do apportionment options track true biomass? Relative percent difference in OM biomass by area vs EM apportionment by area

## For each apportionment option, what is the proportion of years*sims where EM F/F40<1?

## For each apportionment option, what is the proportion of years*sims where EM B/B40<1? What is the mean and median depletion (SSBterminal year/B40) across sims for each year from EM? (plot it)

## mean and median age at harvest from EM (all areas combined) (and roughly mapped to size)
## mean and median catch/ABC for each area and overall

#Proportion years/sims where of the catch in area x is more than 50% (or a changeable value) fish younger than the age @50% mature
#test.age <- 4

#Mean, median (+CIs) of age and length (converted from age?) for catches (apportioned TAC) in each EM area across years and sims


